Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-01-21 src/ms.tex
paper edits
adding discussion of fixed Rocrit fits

\subsection{Initial observations}
A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}. Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}). A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample. As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.

@@

%The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
    \caption{}
    \label{fig:rocrit}
\end{figure*}

>>>>

\subsection{Initial observations}
A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}. Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}). A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample. {\color{red} We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up.} As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.

@@

@@

{\color{red} To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}. The first model assumes \rocrit = 2 and has one free parameter, a temperature shift added to the data. The second model assumes no temperature shift and allows \rocrit to vary (\rocrit being the only free parameter). The results of these fits are shown in Figure~\ref{fig:rocrit}. We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = 2 and variable \rocrit models was negligible. In other words, the \rocrit = 2 model fits the data well with a \teff shift of $\approx$~80--150~K depending on the \teff source used. For the \lamostmcq sample, there is weak support for the variable \rocrit model. However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS). Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure. Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs. \lamostmcq samples (see Fig.~\ref{fig:percentiles}).}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
    \caption{Non-linear least-squares fits to the long-period pile-up for a \rocrit~=~2 model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue). The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=2} - \chi^2_{\Delta T_{eff}=0}$.}
    \label{fig:rocrit}
\end{figure*}
